% Chapter Template
\chapter{Audio Processing} % Main chapter title

\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
Before we feed the audio clip to our machine learning model, it is crucial to pre-process the signal as to achieve higher accuracy
and avoid further deterioration.
The choice and implementation of noise filter will then be explained in \textbf{LINK TO SECTION 1}. We then feed the filtered output 
to a pitch detection algorithm (PDA) \textbf{lINK TO SECTION 2} and then a key detection algorithm (KDA) \textbf{LINK TO SECTION 3}
Figure \textbf{UPDATE!} shows the flowchart of the audio processing part of our project.

% \begin{figure}
% \tikzset{
% every edge/.style = {draw=cyan, line width=2mm, shorten >=1pt, shorten <=1pt,
%                      -{Triangle[scale=0.6]}},
% N/.style args = {#1/#2}{fill=#1, text width=#2, 
%                         font=\scriptsize, align=center},
%    N/.default = cyan/7em,
% }
% \newcommand\tn[1]{\textbf{\small #1}}
%     \begin{tikzpicture}[node distance=8mm]
% \node (n1) [N]  {\tn{User sings into our app}\\ 
%                  (Obtain input audio signal)}; 
% \node (n2) [N=cyan/8em, right=of n1] 
%                 {\tn{Trim silence at the start/end of audio clip};
% \node (n3) [N, right=of n2] 
%                 {\tn{Implement noise filter}};
% \node (n4) [N, below=of n3]
%                 {\tn{Spectral reduction}};
% \node (n5) [N, right=of n4]
%                 {\tn{Low-pass filter}};
% \node (n6) [N, right=of n3] 
%                 {\tn{Implement PDA}};
% \node (n7) [N, right=of n6] 
%                 {\tn{Implement KDA}};
% \node (n8) [N=red!30/8em, below=of n7]
%                 {\tn{Pass the output to Ed's ML model}\\
% 				(Notes and keys)};

% \draw   (n1) edge (n2)
%         (n2) edge (n3)
%         (n3) edge (n4)
% 		(n4) edge (n5)
%         (n3) edge (n6)
%         (n6) edge (n7)
% 		(n7) edge (n8));
%     \end{tikzpicture}
% \end{figure}
%----------------------------------------------------------------------------------------
%	SECTION 0
%----------------------------------------------------------------------------------------
\section{Assumptions}
Before we delineate the approach to audio processing, there are some assumptions that our model
is built on:
\begin{itemize}
	\item \textbf{Assumption 1:} Users' audio input device does not contain active noise cancelling functions.
\end{itemize}

These assumptions will be referred to later on in the section.
%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Noise Filter}
Noise filtering is essential as it reducess or eliminates the noise present in the input signal.
A conventional method to quantify noise is to use signal-to-noise ratio (SNR), which is often 
represented in decibels.
\[SNR=10*log_10((P_{signal})/(P_{noise}))\]
As its name suggests, SNR is the power ratio between desired signal and undesired noise. Effectively,
we would like to use noise filters to achieve a higher SNR.\\ 
There is a few sources of noise when an user record himself with a microphone.
Firstly, there exists self-noise, which is the instrument noise produced by the microphone itself.
Noise may be induced or created when the signal passes through electronic componenets like transistors 
and printed circuit boards.\cite{selfnoise} 
The second source, ambient noise, contributes to a large portion of noise present in a recording.
Room reflections, extraneous noise, electromagnetic interference and mechanical noise are some causes 
to the existence of ambient noise. 
%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\subsection{Possible Models}
Most of the noise filters work in the frequency and spectral domain, 

Spectral-domain models are not suitable for the estimation of extreme responses
\begin{enumerate}
	\item Low-pass filter (LPF)\\
	LPF passes signals with \(f<f_{c}\), where \(f_{c}\) is the cut-off frequency, and attenuates
	signals with \(f>fc\). 
	\[H(f) = rect(f/(2*B))\]
	\[h(t))= \mathfrak{F}^-1{H(f)} = \int_-B^B e^(2*pi*i*f*t)\,df = 2*B*sinc(2*B*t)\]
	In order to implement an LPF, we have to transform signal from time domain to 
	frequency domain using fourier transform. An ideal LPF would completely remove frequencies that are
	higher than \(f_{c}\) and is a non-ca)sual linear time-invariant system. The impulse
	response of an LPF is a sinc function that extends to [$\infty$,-$\infty$]. This is why it is impossible to 
	realize an ideal LPF since that will take infinite time and memory.\\
	LPF avoids aliasing since it removes the high-frequency content but not the desired signal

	\item Wavelet transform\\
	Wavelet transform creates a representation of the signal in both time and frequency domain so localized 
	information of the signal can be efficiently accessed. It is often compared with fourier transform (FT), which
	has the below limitations: 
	\begin{enumerate}
		\item For windowed FT, if the feature is larger or shorter than the window, it cannot be captured completely.
		\item Time resolution for high frequencies is the same for low frequencies. As frequency increases, rate of 
		change of the signal increases, and high frequency signals contain more information in a window than that of 
		low frequency, thus we need a higher time resolution for that.
	\end{enumerate}
	A wavelet is a function that divides the signal into different scale components which 
	
	
	It is more appropriate to describe the process of using wavelet transform to remove noise as a decomposition,rather 
	than a filter.
	Usually for stationary signals, we use conventional frequency-based filters, but for non-stationary ones, we can 
	use wavelet transform, in particular continuous wavelet transform (CWT) which gives better time-scale information
	compared to short-time fourier transform (STFT). 
 
	Wavelet transform analyses a signal into different frequencies at different resolution (multiresolution analysis). 

Advantages of DWT:
	The wavelet expansion allows a more accurate local description and separation of signal characteristics. A 
	Fourier coefficient represents a component that lasts for all time and, therefore, temporary events must be 
	described by a phase characteristic that allows cancellation or reinforcement over large time periods. 
	A wavelet expansion coefficient represents a component that is itself local and is easier to interpret. 
	The wavelet expansion may allow a separation of components of a signal whose Fourier description overlap in both time and frequency.
	
	Wavelets are adjustable and adaptable. Because there is not just one wavelet, they can be designed to fit 
	individual applications. They are ideal for adaptive systems that adjust themselves to suit the signal.
Disadvantages of DWT:
	Sensitive to shifting
	Poor directionality
	Lack of phase information
	Hard to choose an appropriate mother wavelet and number of decomposition levels
s

D. Spectral reduction
Define our noisy audio y(n) = x(n) + d(n), for 0<= n <= N-1, where x(n) is our original signal (signal we wish to recover), d(n) is the noise, n is the time index, N is the number of samples. 
Assuming d(n) and x(n) have no correlation, Y(omega)= 
Used by audacity
https://github.com/timsainb/noisereduce
	Did a really good job, able to filter white noise, but not sudden background noise (I accidentally used an audio with a sharp squeal in the beginning and it didn’t get filtered, which makes sense if we think about spectral subtraction, it is impossible to subtract an unexpected frequency)
	Will use spectral reduction in our project!!! 
\end{enumerate}
%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\subsection{Implementation}
LPF: Cutoff frequency selection: depending on the gender and age, chebyshev? elliptical?
\begin{table}[]
	\begin{tabular}{lll}
																				 & Male   & Female  \\
	Pitch range (Hz)                                                             & 60-180 & 160-300 \\
	\begin{tabular}[c]{@{}l@{}}Praat pitch parameters \\ range (Hz)\end{tabular} & 50-300 & 100-600
	\end{tabular}
\end{table}
%-----------------------------------
%	SUBSECTION 3
%-----------------------------------

\subsection{Improvements}
use the silence to feed in spectral reduction

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

\section{Pitch Detection Algorithm (PDA)}


%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\subsection{Possible Models}

%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\subsection{Implementation}
Morbi rutrum odio eget arcu adipiscing sodales. Aenean et purus a est pulvinar pellentesque. Cras in elit neque, quis varius elit. Phasellus fringilla, nibh eu tempus venenatis, dolor elit posuere quam, quis adipiscing urna leo nec orci. Sed nec nulla auctor odio aliquet consequat. Ut nec nulla in ante ullamcorper aliquam at sed dolor. Phasellus fermentum magna in augue gravida cursus. Cras sed pretium lorem. Pellentesque eget ornare odio. Proin accumsan, massa viverra cursus pharetra, ipsum nisi lobortis velit, a malesuada dolor lorem eu neque.

%----------------------------------------------------------------------------------------
%	SECTION 3
%----------------------------------------------------------------------------------------
\section{Key Detection Algorithm (KDA)}

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\subsection{Possible Models}

%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\subsection{Implementation}
